---
title: "Mini project"
author: "Alicia Pacheco"
date: "October 30, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the data 

```{r}
url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(url)
#how many patients
nrow(wisc.df)

#diagnostics

table(wisc.df$diagnosis)


```



Diagnostics

```{r}
#cheking for malignant..
diagnosis<-as.numeric(wisc.df$diagnosis=="M")

```





Convert the features of the data: wisc.data

```{r}

#make a new matrix from colums 3 to 32 in wisc.df  
wisc.data <- as.matrix(wisc.df[3:32] )

#
id<-colnames(wisc.data)

# Set the row names of wisc.data  

row.names(wisc.data) <- wisc.df$id
#head(wisc.data)
```
Variables/features in the data are suffixed with _mean

```{r}
#grep() looks for patterns 

x<-length(grep("_mean",colnames(wisc.data)))


```

There are `r x` mean measurements in this dataset

##Performing PCA (min58)


```{r}

# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)

```


```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data,scale=TRUE)
summary(wisc.pr)
#attributes
attributes(wisc.pr)
dim(wisc.pr$x)


```

#Interpreting PCA results

```{r}
biplot(wisc.pr,scale=TRUE)
```


```{r}
#We want to plot PC1 and PC2, which is wisc.pr$x[,1] and wisc.pr$x[,2] respectively


plot(wisc.pr$x[,1],wisc.pr$x[,2],col=diagnostics+1,xlab = "PC1", ylab = "PC2")

    
```


```{r}
# Repeat for components 1 and 3
#code is different but gives same result as code for PC1 and PC2
plot(wisc.pr$x[, c(1, 3)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")


```

Because principal component 2 explains more variance in the original data than principal component 3, you can see that the first plot has a cleaner cut separating the two subgroups.  

Variance explained

start studying here.......

```{r}
pr.var<-wisc.pr$sdev^2

# Variance explained by each principal component: pve
pve <-  pr.var/sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

Alternative data plot

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )


```

```{r}
# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```



##Section 3

Scale the data

```{r}
# Scale the wisc.data data: data.scaled
#data.scaled <- hclust(wisc.data,)

data.scaled<-scale(wisc.data)
data.dist<-dist(data.scaled)
wisc.hclust<-hclust(data.dist)

plot(wisc.hclust)
wisc.hclust.clusters<-cutree(wisc.hclust,k=4)
table(wisc.hclust.clusters)
table(wisc.hclust.clusters,diagnostics)

```

selecting numbers of clusters

```{r}




```

##Section 5

Clustering on PCA results

```{r}
# Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
#wisc.pr$x[ ,1:7]
dis.pr<-dist(wisc.pr$x[ ,1:7])
wisc.pr.hclust <- hclust(dis.pr, method="complete")

wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
table(wisc.hclust.clusters)
table(wisc.hclust.clusters,diagnostics)

```

## Bonus: Predicting modeling with PCA Components


```{r}

## Predicting Malignancy Of New samples
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
plot(wisc.pr$x[,1:2], col=diagnostics+1)
points(npc[,1], npc[,2], col="blue", pch=16, cex=2)
#patient near red=bad, patient near black=good

```

