---
title: "Class08 try at home"
author: "Alicia Pacheco"
date: "October 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering 
Our first example with k-means 
```{r}
#Example plot see how R markdown works
plot(1:10,type="l")

#generate some example data for clustering 
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))  
#And we plot
plot(x)


```
  
Using the *kmeans()* function setting k to 2 and nstart=20
Inspect/print the results

```{r}
k<-kmeans(x,centers=2,nstart=20)

```
How mant points are in each cluster?
```{r}
k$size

```
 Q. What ‘component’ of your result object details
 - cluster size
  `k$size`  
 - cluster assignment/membership
  `k$cluster`  
 - cluster center?
  `k$centers`
  
```{r}

#cluster size
k$size



```

```{r}
#cluster assignement/membership
k$cluster


```

```{r}
#cluster centers
k$centers
```
Plot x colored by the kmeans cluster assignment and
add cluster centers as blue points

```{r}
plot(x, col=k$cluster,pch=20)
palette(c("black","red"))
points(k$centers,col="blue3",pch=20,cex=3)

```

Q. Repeat for k=3, which has the lower tot.withinss? 
 
```{r}
k3<-kmeans(x,centers=3, nstart = 20)

plot(x,col=k3$cluster,pch=20)
palette(c("purple","red","black"))
points(k3$centers,col="blue",pch=20,cex=2)


```
 
```{r}
#total within cluster sum of squares, the smallest the petter 
k$tot.withinss

k3$tot.withinss
```
 
##Hierarchical CLustering  
Create a hierarchical cluster model:hc
```{r}
hc<-hclust(dist(x))
plot(hc)
```
cutree

```{r}
  
cutree(hc,k=2)

```
###A more real example

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))

colnames(x) <- c("x", "y")
```

```{r}
# Step 2. Plot the data without clustering
plot(x)

```
```{r}

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

Q. Use the dist(), hclust(), plot() and cutree()
functions to return 2 and 3 clusters 

```{r}

hc<-hclust(dist(x))

plot(hc)
cutree(hc,k=2)  


```

```{r}
hc<-hclust(dist(x))

plot(hc)
cutree(hc,k=3)
```

##**PCA**

First lets generate some example data to work with. 

```{r}
 #Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10) 
#Labeled the rows gene1, gene2, etc all the way to gene100  

rownames(mydata)<-paste("gene",1:100,sep="")
#Label the columns with wt and ko  
colnames(mydata)<-c(paste("wt",1:5,sep=""),paste("ko",1:5,sep=""))
head(mydata)

```
Fill in some facke read counts

```{r}
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
} 


```

Now we can do PCA  


```{r}
#prcomp expects the samples to be the row and genes to be columns so we need to transose the matrix with the t()  
pca<-prcomp(t(mydata),scale=TRUE)

#See atributes  
attributes(pca)

```
pca$ contains the principal components

```{r}

plot(pca$x[,1],pca$x[,2])
```

variance captured per Pc

```{r}
#variance captured by PCA

pca.var <- pca$sdev^2
# Precent variance is often more informative to look at

pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 

pca.var.per

```
We can plot this 

```{r}
barplot(pca.var.per, main="Scree Plot", xlab="Principal Component", ylab="Percent Variation")
#This means that there are big differences between these two groups

```


We can add color to the vector 

```{r}
colvec<-colnames(mydata)
colvec[grep("wt",colvec)]<-"red"
colvec[grep("ko",colvec)]<-"blue"

plot(pca$x[,1],pca$x[,2],col=colvec, pch=20,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

```

We can identify the points

```{r}
plot(pca$x[,1],pca$x[,2],col=colvec, pch=20,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

identify(pca$x[,1],pca$x[,2],labels=colnames(mydata))

```

Finally we will use loading scores

```{r}
#loading scores help us deteremine which genes have the largest effect on where samples are pltted in the PCA plot
#prcomp() calls loading scores $rotation

loading_scores<-pca$rotation[,1]

summary(loading_scores)

```
We are interested in the magnitued of both plu and minus contributing genes

```{r}
gene_scores<-abs(loading_scores)

gene_scores

```

```{r}
#we can sort magnitudes from high to low  

gene_score_ranked <- sort(gene_scores, decreasing=TRUE)  



```
```{r}
#lets se the top five scores  
top_five<-names(gene_score_ranked[1:5])  
top_five


```

```{r}
##show the scores

pca$rotation[top_five,1]  

```


