---
title: "Class14"
author: "Alicia Pacheco"
date: "November 15, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Class 14 Examine Astma SNP's 

##Section 1: Identify genetic variants of interest

Asthma SNPs in MXL ansestry
```{r}

mxl<-read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(mxl)
```

Now focus on 2nd column that contains geneotype

```{r}

#genotype percentages

genotypes<-round(table (mxl[,2])/nrow(mxl)*100,2)

genotypes

```


## Section 2: Initial RNA-seq anlysis  

4th line in the fastq files are the scores in ascii
```{r}

#install.packages("gtools")
#install.packages("seqinr")


library(seqinr)
library(gtools) 
phred <- asc(s2c("DDDDCDEDCDDDDBBDDDCC@") ) - 33
phred 

```


##Section 4: Population scale analysis 

One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale.  
Read file into project  


```{r}
#read the file 
expr<-read.table("rs8067378_ENSG00000172057.6.txt")

inds.aa<-expr$geno=="A/A"


summary(expr$exp[inds.aa])
```

```{r}
#This is the variation 
inds.ag<-expr$geno=="A/G"


summary(expr$exp[inds.ag])
```
```{r}
#This is teh WT
inds.gg<-expr$geno=="G/G"


summary(expr$exp[inds.gg])
```
We can make a boxplot here

```{r}
#by looking at the Insect Spray example we conclude that

boxplot(exp ~ geno, data = expr)

```

Does the SNP effect the expression of ORMDL3?  Yes the AA genotype affects expression of the gene.  
